---
title: "What time of year and destinations are the worst for cancellation?"
author: "Crystal Tse"
date: "2/8/2020"
output: html_document
---

```{r setup, include = FALSE}
library(mosaic)
library(tidyverse)
library(RColorBrewer)

abia = read.csv("abia.csv")
```
When planning travel from Austin, especially with connecting flights, consider cancellations and delays. This analysis shows the months and destinations which are the worst.

```{r echo = FALSE}
# filter out flights that are not cancelled
cancelled = abia %>%
  filter(Cancelled == 1)
```

```{r echo = FALSE}
# group by destination and then summarize with count
# top destinations
ten_most_visited = abia %>% filter(Dest != "AUS") %>% group_by(Dest) %>% summarize(count = n()) %>% arrange(-count)

ten_most_visited = ten_most_visited[1:10,]
```

```{r echo = FALSE}
most_visited = abia[abia$Dest == ten_most_visited$Dest,]
ten_most_visited = inner_join(abia, ten_most_visited, by = "Dest")
```
The graph below plots month and cancellation rate (out of all flights) for the top ten most frequented destinations out of the Austin-Bergstrom International Airport. 

```{r echo = FALSE}
ggplot(data = ten_most_visited, 
  aes(x = Month, y = (Cancelled /(sum(Cancelled == 1))), fill = Dest))+
  geom_bar(stat = "identity") + 
  scale_fill_brewer(palette = "Spectral") + 
  scale_x_discrete(limits = month.abb) +
  ggtitle("Cancellation Rates for Top 10 Destinations") +
  labs(y = "Cancellation Rate", x = "Month") +
  labs(fill = "Destination")

```

One particularly surprising finding is that the largest proportion of cancelled flights occur in March and April, whereas October and November have very low cancellation rates. 

Additionally, in most months, Dallas's major airports (DFW and DAL) experience a large fraction of cancellations, but in September, Houston's two major airports (HOU and IAH) do.

To investigate this surprising pattern of cancellations, it is worth looking into delays for each month.

```{r echo = FALSE}
# Investigate median departure delays by month.

departure_delays_by_month <- abia %>% 
  group_by(Month) %>% 
  filter(DepDelay > 0) %>% 
  # Take the median since the distribution is highly skewed.
  summarize(delay = median(DepDelay, na.rm = TRUE)) %>%
  select(delay) %>% 
  mutate(month = month.name)

ggplot(data = departure_delays_by_month) +
  geom_point(
    mapping = aes(x = delay, y = month),
    color = "red",
    size = 3
  ) +
  geom_vline(xintercept = 0, size = .25) +
  xlim(c(0, 20)) +
  scale_y_discrete(limits = rev(month.name)) +
  labs(title = "Median Departure Delay by Month", y = "", x = "Delay in Minutes")
```

Departure delays are also very low in October, but in contrast to the high number of cancellations in September, there are relatively few delays. December is worst for departure delays, alongside February and March.

So, February and March are consistently bad for both cancellations and departure delays. December exhibits low cancellation rates but the highest departure delay times.

That begs the question of whether arrival delays exhibit similar patterns.

```{r echo = FALSE}
arrival_delays_by_month <- abia %>% 
  group_by(Month) %>% 
  filter(ArrDelay > 0) %>% 
  # Take the median since the distribution is highly skewed.
  summarize(delay = median(ArrDelay, na.rm = TRUE)) %>%
  select(delay) %>% 
  mutate(month = month.name)

ggplot(data = arrival_delays_by_month) +
  geom_point(
    mapping = aes(x = delay, y = month),
    color = "red",
    size = 3
  ) +
  geom_vline(xintercept = 0, size = .25) +
  xlim(c(0, 20)) +
  scale_y_discrete(limits = rev(month.name)) +
  labs(title = "Median Arrival Delay by Month", y = "", x = "Delay in Minutes")

```

September through November experience relatively few arrival delays, and February and March still experience the greatest arrival delay time; however, the summer months experience more arrival delays than departure delays.

December consistently has the greatest delay time, for both arrivals and departures.

In sum, anticipate a greater chance of cancellations and delays in February and March, and delays in December. Since the highest cancellation rates for top destinations from the Austin-Bergstrom International Airport are within the state, it is worth considering  more reliable transportation options to guarantee reaching your destination.