---
title: SDS 323 Exercises 1
author: Kyle Carter, Jacob Rachiele, Crystal Tse, Jinfang Yan
date: 2/14/2020
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE}
# load required packages
library(mosaic)
library(tidyverse)
```
## Problem 1: Flights at ABIA

Let's investigate the median departure delays by month. We look at the
_median_ since the data is highly skewed, as you see here.

```{r}
# Read in the Austin 2008 flights data

abia <- read.csv("./data/ABIA.csv", header = TRUE)

ggplot(data = abia) +
  geom_histogram(mapping = aes(x = DepDelay),
                 na.rm = TRUE, binwidth = 20) +
  xlim(0, 500) + ylim(0, 12500)
```

The following figure shows the median departure delay by month.

```{r}
departure_delays_by_month <- abia %>% 
  group_by(Month) %>% 
  filter(DepDelay > 0) %>% 
  # Take the median since the distribution is highly skewed.
  summarize(delay = median(DepDelay, na.rm = TRUE)) %>%
  select(delay) %>% 
  mutate(month = month.name)


ggplot(data = departure_delays_by_month) +
  geom_point(
    mapping = aes(x = delay, y = month),
    color = "red",
    size = 3,
    alpha = 0.6
  ) +
  geom_vline(xintercept = 0, size = .25) +
  xlim(c(0, 20)) +
  scale_y_discrete(limits = rev(month.name)) +
  labs(title = "Median Departure Delay by Month", y = "", x = "Delay in Minutes")
```

Here we note some suprising data. December has a high median departure delay as expected, but March has a higher median delay than some the winter months of November,
January, and February. Maybe looking at arrival delays could shed some light on whether
the delay pattern by month is because of seasonality surrounding Austin.

```{r}
arrival_delays_by_month <- abia %>% 
  group_by(Month) %>% 
  filter(ArrDelay > 0) %>% 
  # Take the median since the distribution is highly skewed.
  summarize(delay = median(ArrDelay, na.rm = TRUE)) %>%
  select(delay) %>% 
  mutate(month = month.name)

ggplot(data = arrival_delays_by_month) +
  geom_point(
    mapping = aes(x = delay, y = month),
    color = "red",
    size = 3,
    alpha = 0.6
  ) +
  geom_vline(xintercept = 0, size = .25) +
  xlim(c(0, 20)) +
  scale_y_discrete(limits = rev(month.name)) +
  labs(title = "Median Arrival Delay by Month", y = "", x = "Delay in Minutes")
```

Once again, we note a similar pattern -- arrival delays in March (as well as June)
are suprisingly high relative to the median delays in some of the winter months.

## Problem 2: Regression Practice (Creatinine)

## Problem 3: Green Buildings

## Problem 4: Milk Prices

```{r}
# Read the data file
milk <- read.csv("./data/milk.csv")
```

First, graph the data.

```{r message = FALSE}
ggplot(data = milk) +
  geom_point(aes(x = price, y = sales)) +
  geom_smooth(mapping = aes(x = price, y = sales))
```

Notice that this is not a linear relationship, which makes sense since
quantity demanded is modeled in microeconomics using a Power Law:
$Q = KP^E$,
where Q is the quantity demanded, P is the price, E is the price elasticisty of demand and K is a constant.

Step 1: Write an equation that expresses net profit N in terms of both Q and P (and cost c)

$$N = (P - c)Q$$

Step 2: Use the microeconomic model of quanity demanded, which is a function of the price.

$Q = f(P) = KP^E$, so that $N = (P - c)f(P) = (P-c)(KP^E)$

The values of K and E are uknown, so we must estimate them from the data.

We can do this using linear regression using the product and power rules of logarithms, which tell us that $ln(Q) = ln(KP^E) = ln(K) + E(ln(P))$.

This has the form of a simple linear regression, where $\beta_0 = ln(K)$ and 
$\beta_1 = E$.

Step 3: Use simple linear regression to estimate the unknown coefficients.

```{r}
model <- lm(log(sales) ~ log(price), data = milk)
```

Confirm the linearity of the logarithm of the data by plotting.

```{r}
plot(log(sales) ~ log(price), data = milk)
abline(model, col = "red")
```

Now we have an estimate of ln(K) in the form of the intercept of the model, `r round(coef(model)[1], 2)`, and of E in the form of the slope of the model, `r round(coef(model)[2], 2)`

```{r echo = FALSE}
K <- exp(coef(model)[1])
E <- coef(model)[2]
```

Taking the exponential of both sides gives us net profit in terms of P and c alone,
$N \approx  (P - c)(112P^{-1.62})$

Let's assume c = 1.
```{r}
x <- milk$price
c <- 1
curve((x - c) * K * x^(E), from = 1, to = 9, xlab = "Price", ylab = "Sales")
#Zoom in 
curve((x - c) * K * x^(E), from = 2.5, to = 2.7, xlab = "Price", ylab = "Sales")
#Zoom in more
curve((x - c) * K * x^(E), from = 2.60, to = 2.63, xlab = "Price", ylab = "Sales")
```

From the final plot, we see that the price that maximizes net profit is close to $2.62.
