---
title: SDS 323 Exercises 1
output: md_document
---

## SDS 323 Exercises 1
### Kyle Carter, Jacob Rachiele, Crystal Tse, Jinfang Yan
### 2/14/2020

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
# load required packages
library(mosaic)
library(tidyverse)
```
## Problem 1: Flights at ABIA

## Problem 2: Regression Practice (Creatinine)



## Problem 3: Green Buildings

## Problem 4: Milk Prices

```{r}
# Read the data file
milk <- read.csv("data/milk.csv")
```

First, graph the data.

```{r message = FALSE}
ggplot(data = milk) +
  geom_point(aes(x = price, y = sales)) +
  geom_smooth(mapping = aes(x = price, y = sales))
```

Notice that this is not a linear relationship, which makes sense since
quantity demanded is modeled in microeconomics using a Power Law:
$Q = KP^E$,
where Q is the quantity demanded, P is the price, E is the price elasticisty of demand and K is a constant.

Step 1: Write an equation that expresses net profit N in terms of both Q and P (and cost c)

$N = (P - c)Q$

Step 2: Use the microeconomic model of quanity demanded, which is a function of the price.

$Q = f(P) = KP^E$, so that $N = (P - c)f(P) = (P-c)(KP^E)$

The values of K and E are uknown, so we must estimate them from the data.

We can do this using linear regression using the product and power rules of logarithms, which tell us that $ln(Q) = ln(KP^E) = ln(K) + E(ln(P))$.

This has the form of a simple linear regression, where $\beta_0 = ln(K)$ and 
$\beta_1 = E$.

Step 3: Use simple linear regression to estimate the unknown coefficients.

```{r}
model <- lm(log(sales) ~ log(price), data = milk)
```

Confirm the linearity of the logarithm the data by plotting.

```{r}
plot(log(sales) ~ log(price), data = milk)
abline(model, col = "red")
```

Now we have an estimate of ln(K) in the form of the intercept of the model, `r round(coef(model)[1], 2)`, and of E in the form of the slope of the model, `r round(coef(model)[2], 2)`

```{r echo = FALSE}
K <- exp(coef(model)[1])
E <- coef(model)[2]
```

Taking the exponential of both sides gives us net profit in terms of P and c alone,
$N \approx  (P - c)(112P^{-1.62})$

Let's assume c = 1.
```{r}
x <- milk$price
c <- 1
curve((x - c) * K * x^(E), from = 1, to = 9, xlab = "Price", ylab = "Sales")
#Zoom in 
curve((x - c) * K * x^(E), from = 2.5, to = 2.7, xlab = "Price", ylab = "Sales")
#Zoom in more
curve((x - c) * K * x^(E), from = 2.60, to = 2.63, xlab = "Price", ylab = "Sales")
```

From the final plot, we see that the price that maximizes net profit is close to $2.62.
